### 2020-02-27 ChapterBackground/chapterBackground.tex
Updated: Latency and Packet-loss values in Background re-write

This study is concerned with inter-region and inter-AZ latencies as these are the latencies that affect server-server communication and the correctness of the system. Lower latency is ideal for the deployment of a real-time distributed physics system, in the case of AP, lower latency means smaller auras leading to better partitioning and subsequent performance. However, inter-AZ deployment may not always be possible or desired (e.g. lack of availability of resources in one AZ or to reduce latency between users and some of the servers in the system in the case of geographically diverse users). Therefore the latency experiments used Inter-AZ latency and low Inter-region latency values.  The lowest latency observed was from the inter-AZ category, i.e. servers hosted in the same geographical region at $<2ms$. Inter-region latency measured in the range of ($7-302ms$). The chosen values for the latency experiment were as follows: $2ms$, $8ms$ and $16ms$. These values represent inter-AZ latency and low inter-region latencies within AWS, for regions that lie within same continent (eu-West-2 - eu-west-3) and (eu-West-1 - eu-west-2) \cite{ThousandEyesCloudPerf2019}.

%Dale's background starts here
\subsection{Packet Loss Values}

Packet loss is the condition where data travelling across a computer network fails to arrive at the destination. The most common reasons for packet loss include network congestion, errors in transmission (e.g. in wireless networks), or hardware failure. According to a study carried out in 2018, packet loss within cloud providers' networks is negligible (measured at 0.01\% on average), even when traversing inter-region backbone links, or when traffic is being sent between different cloud providers, indicating a very high level of reliability \cite{ThousandEyesCloudPerf2018}. It was found that packet loss only became significant when measured between certain geographical ``end-user'' locations outside of the cloud providers' networks, and even so, was rarely measured above 1\%, with the worst case being traffic travelling in or out of China, irrespective of the cloud provider (6.3\%).

We therefore do not expect a typical cloud deployment of a distributed physics system to be subject to significant packet loss, although we decided to test our system's behaviour up to what would be considered an extreme level packet loss to see what kind of effect the loss of data would have on the accuracy of the simulation. A maximum value of 20\% was used to ensure that any kind of trend would be visible.

%Dale's background ends here

>>>>

This study is concerned with inter-region and inter-AZ latencies as these are the latencies that affect server-server communication and the correctness of the system. Lower latency is ideal for the deployment of a real-time distributed physics system, in the case of AP, lower latency means smaller auras leading to better partitioning and subsequent performance. However, inter-AZ deployment may not always be possible or desired (e.g. lack of availability of resources in one AZ or to reduce latency between users and some of the servers in the system in the case of geographically diverse users). Therefore the latency experiments used Inter-AZ latency and low Inter-region latency values.  The lowest latency observed was from the inter-AZ category, i.e. servers hosted in the same geographical region at $<2ms$. Inter-region latency measured in the range of ($7-302ms$).

The chosen values for the latency experiment were as follows: $2ms$, $8ms$ and $16ms$. These values represent inter-AZ latency and low inter-region latencies within AWS, for regions that lie within same continent (eu-West-2 - eu-west-3) and (eu-West-1 - eu-west-2) \cite{ThousandEyesCloudPerf2019}.

%Dale's background starts here
\subsection{Packet-Loss Values}
Another factor that affects the performance of networked applications is packet-loss. An experiment was conducted in this study to determine the effects of packet-loss on the correctness of AP. The most common causes of packet-loss include network congestion \cite{1998PacketLoss} and bit errors in wireless transmission \cite{2005PacketLoss}. According to \cite{ThousandEyesCloudPerf2018}, carried out in 2018, cloud providers' networks have a very high level of reliability and exhibit only negligible packet-loss (measured at 0.01\% of packets lost, on average). This holds true even when traversing inter-region backbone links and between different cloud providers. The only significant measured level of packet-loss was between some geographical ``end-user'' locations outside of the cloud providers' networks, and were typically $<1\%$. However, this study is only concerned with packet-loss between servers in the cloud and not ``end-users''. The only notable exception to this and worst case for packet-loss was for traffic entering or exiting China, which experienced packet-loss as high as $8\%$.

The chosen values for the packet-loss correctness experiment range from $0\%-20\%$. These values cover the entire range of expected packet-loss values while running in the cloud, as well as beyond those values, in order to determine what impact severe packet-loss would have on the correctness of AP and establish the trend increasing packet-loss has.
